NN_Accel_Heading

### 2024-03-07 Fish Data Analysis/NN_Accel_Heading.qmd
Modeling angular data

@@

heading_ps_angle_base <- lm(Fish1_Yaw_Heading_PS ~ 1, data = fish_data_nn %>% filter(Flow == "Still Water"))

heading_ps_angle_step <- stepAIC(heading_ps_angle_base, 
                                   list(upper = ~ (Angle_Rads+cos(Angle_Rads)+sin(Angle_Rads))*Ablation*Darkness, lower = ~1), 
                                   direction = "both",
                                   trace = F)

@@

ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Angle_Rads, y = Fish1_Yaw_Heading_PS, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  theme_classic() +
  #ylim(-5,5)+
  xlab("direction of neighbour (rad)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")

ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = mean_turn, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  geom_line(alpha = 0.75) +
  #geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  theme_classic() +
  xlab("direction of neighbour (rad)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")

>>>>

Estimate the Wavelength

```{r}
# function to be minimized
sum_squares <- function(lambda) {
  lmfit <- lm(Fish1_Yaw_Heading_PS ~ (sin(2*pi/lambda*Fish1_Yaw_Heading_PS) + cos(2*pi/lambda*Fish1_Yaw_Heading_PS))*Ablation*Darkness,
              data=fish_data_nn %>% filter(Flow == "Still Water"))
  return(sum(lmfit$residuals^2))
}

lambda <- optimize(sum_squares,c(1,100))$minimum
lambda

#par 1 = amp
#par 2 = freq
#par 3 = phase

min.RSS <- function(data, par) {
             with(data, sum(par[1]*sin(par[2]*Angle_Rads+par[3]) - Fish1_Yaw_Heading_PS)^2)
}

result_F0LNDN <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Light"))
result_F0LYDN <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, Ablated, Light"))
result_F0LNDY <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Dark"))
result_F0LYDY <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, Ablated, Dark"))

result_F0LNDN$par
result_F0LYDN$par
result_F0LNDY$par
result_F0LYDY$par

#par 4 & 5 = amp Ablation & Darkness
#par 6 & 7 = freq Ablation & Darkness
#par 8 & 9 = phase Ablation & Darkness

fish_data_nn_optim <- fish_data_nn %>% filter(Flow == "Still Water") %>% mutate(AblationCat = ifelse(Ablation == "Ablated",1,0),
                                                                                DarknessCat = ifelse(Darkness == "Dark",1,0))

min.RSS_cat <- function(data, par) {
             with(data, sum((par[1]+par[4]*AblationCat+par[5]*DarknessCat)*
                              sin((par[2]+par[6]*AblationCat+par[7]*DarknessCat)*Angle_Rads+
                                    (par[3]+par[8]*AblationCat+par[9]*DarknessCat)) - Fish1_Yaw_Heading_PS)^2)
}

result_F0_cat <- optim(par = c(1, 1, 0, 0, 0, 0, 0, 0, 0), fn = min.RSS_cat, data = fish_data_nn_optim)

result_F0_cat

nn_heading_sum <- nn_heading_sum %>% mutate(AblationCat = ifelse(Ablation == "Ablated",1,0),
                                            DarknessCat = ifelse(Darkness == "Dark",1,0)) %>%
                                     mutate(sin_pred = (result_F0_cat$par[1]+result_F0_cat$par[4]*AblationCat+result_F0_cat$par[5]*DarknessCat)*
                                                        sin((result_F0_cat$par[2]+result_F0_cat$par[6]*AblationCat+result_F0_cat$par[7]*DarknessCat)*angle_bins+
                                                        (result_F0_cat$par[3]+result_F0_cat$par[8]*AblationCat+result_F0_cat$par[9]*DarknessCat)))

# #Now let's check the F0LNDN condition
# heading_ps_F0LNDN_lm_base <- lm(Fish1_Yaw_Heading_PS ~ 1, 
#                                 data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Light"))
# heading_ps_F0LNDN_lm_lin <- lm(Fish1_Yaw_Heading_PS ~ Angle_Rads, 
#                                 data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Light"))
# heading_ps_F0LNDN_lm_sin <- lm(Fish1_Yaw_Heading_PS ~ sin(0.4444551*Angle_Rads+0.2544464), 
#                                 data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Light"))
# 
# AIC(heading_ps_F0LNDN_lm_base,heading_ps_F0LNDN_lm_lin,heading_ps_F0LNDN_lm_sin) %>% arrange(AIC)
# 
# #Now let's check the F0LYDN condition
# heading_ps_F0LYDN_lm_base <- lm(Fish1_Yaw_Heading_PS ~ 1, 
#                                 data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, Ablated, Light"))
# heading_ps_F0LYDN_lm_lin <- lm(Fish1_Yaw_Heading_PS ~ Angle_Rads, 
#                                 data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, Ablated, Light"))
# heading_ps_F0LYDN_lm_sin <- lm(Fish1_Yaw_Heading_PS ~ sin(1.11686351*Angle_Rads+-0.04623624), 
#                                 data = fish_data_nn %>% filter(Flow_Ablation_Darkness == "Still Water, Ablated, Light"))
# 
# AIC(heading_ps_F0LYDN_lm_base,heading_ps_F0LYDN_lm_lin,heading_ps_F0LYDN_lm_sin) %>% arrange(AIC)

@@

fish_data_nn_model <- fish_data_nn %>% filter(Flow == "Still Water") %>% 
                                       mutate(AblationCat = ifelse(Ablation == "Ablated",1,0),
                                              DarknessCat = ifelse(Darkness == "Dark",1,0))

heading_ps_angle_base <- lm(Fish1_Yaw_Heading_PS ~ 1, data = fish_data_nn_model)

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ ((Angle_Rads+cos(Angle_Rads)+sin(Angle_Rads))*Ablation*Darkness), lower = ~1),
                                   direction = "both",
                                   trace = F)

@@

tester_model <- lm(Fish1_Yaw_Heading_PS ~ sin(Angle_Rads):Ablation:Darkness, data = fish_data_nn_model)

ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Angle_Rads, y = Fish1_Yaw_Heading_PS, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F)+
  #geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  #geom_line(aes(y = predict(heading_ps_angle_step)), size = 1)+
  theme_classic() +
  #ylim(-5,5)+
  xlab("direction of neighbour (rad)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")

ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = mean_turn, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  #geom_smooth(se = F)+
  #geom_line(aes(y = sin_pred)) +
  #geom_line(alpha = 0.75) +
  #geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  geom_line(aes(y = predict(heading_ps_angle_step,
                    newdata = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Angle_Rads = angle_bins))), size = 1)+
  theme_classic() +
  xlab("direction of neighbour (rad)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")
